### 2022-01-26 article-counts-per-year-figure.Rmd
renamed file

---
title: "Citations figure"
author: "Fred Boehm & Xiang Zhou"
date: "1/26/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

We need a bar plot of counts per year for a search with keywords "Mendelian randomization".

```{r, eval = FALSE}
remotes::install_github("akshaynagpal/RgScholar")
```

```{r}
RgScholar::google_Scholar(user_query = "mendelian randomization", year_low = 2021)
```