### 2022-01-31 mr-review-outline.Rmd
switched outline to md output format

---
title: "Mendelian randomization for causal inference in Genome-wide association studies"
author: "Frederick J. Boehm"
date: "`r lubridate::now()`"
bibliography: exported.bib
csl: elsevier-with-titles.csl
output: 
  - bookdown::github_document2
  - bookdown::html_document2
  - bookdown::word_document2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

# Mendelian Randomization with one Instrument

## Assumptions of Mendelian Randomization

## Assumption violations

## Statistical models for MR with one instrument and one outcome

## Statistical methods for MR with one instrument and one outcome

# Mendelian Randomization with multiple instruments

# Horizontal pleiotropy

# Correlated horizontal pleiotropy

# Recent Findings

## Omnigenic MR

## Multiple outcomes in MR

# Applications of MR in GWAS

# Summary and Outlook

# References