### 2022-01-31 article-counts-per-year-figure.Rmd
magrittr to replace dplyr

---
title: "Article counts per year figure"
author: "Fred Boehm"
date: "1/26/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

We need a bar plot of counts per year for a search with keywords "Mendelian randomization".

I manually recorded the article counts per year for two searches: 
1. genome-wide association study and 2. mendelian randomization.

```{r}
library(magrittr)
dat <- readxl::read_xlsx(path = "data/publication-counts.xlsx", sheet = "counts")
dat %>%
  ggplot2::ggplot(ggplot2::aes(x = year, y = mrcount)) + ggplot2::geom_bar(stat = "identity")
```

```{r}
dat %>%
  ggplot2::ggplot(ggplot2::aes(x = year, y = gwascount)) + ggplot2::geom_bar(stat = "identity")

```