### 2022-02-27 mr-review.Rmd
added text about MR GENIUS

---
title: "Mendelian randomization for causal inference in Genome-wide association studies"
author: "Frederick J. Boehm & Xiang Zhou"
date: "`r lubridate::now()`"
bibliography: exported.bib
csl: elsevier-with-titles.csl
link-citations: yes
linkcolor: blue
output: 
  - bookdown::pdf_document2
  - bookdown::html_document2
  - bookdown::word_document2
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
library(magrittr)
```

# Abstract

Complex trait genetics studies in humans have yielded thousands of new associations for  
diverse traits. The next step in understanding biology of health and disease is to
identify causal relationships among complex traits. Two-sample Mendelian randomization studies 
in genome-wide association studies can uncover causal relationships 
between complex traits. By leveraging shared SNP genotypes as instrumental variables, or proxies, for 
the exposure complex trait, investigators can tease out causal effects from observational data, provided
that necessary assumptions are satisfied. We discuss below the growth of genome-wide association studies 
in parallel with development of Mendelian randomization methods. We argue that the recent availability of GWAS summary statistics for diverse complex traits has motivated new Mendelian randomization methods and that this area continues to offer opportunities for additional methods developments.

# Introduction

Great interest has developed in inferring causal relationships between complex 
traits, *i.e.*, traits that seemingly are not inherited in a Mendelian fashion, in observational human genetics studies. Discovery of such relationships is crucial to 
enhancing our understanding of the biology of health and disease. Mendelian randomization (MR) studies are a form of instrumental variable analysis in which
genetic markers, typically single nucleotide polymorphisms (SNPs), serve as instruments, or proxies, for 
inferring causal effects of an exposure variable on an outcome variable [@wright1928tariff;@wright1921correlation;@wright1923theory;@stock2003retrospectives;@katan1986apolipoprotein;@katan2004commentary;@didelez2007mendelian].

They are termed 
"two-sample" MR studies because the exposure variable and the outcome variable are measured on 
two non-overlapping cohorts of subjects [@pierce_efficient_2013]. Importantly, the SNP instruments are measured in both sets of subjects.

Genome-wide association studies (GWAS) present unique opportunities for 
discovery of causal relationships via MR. A GWAS
interrogates millions of single nucleotide polymorphisms (SNPs) to infer which affect the trait [@the_wellcome_trust_case_control_consortium_genome-wide_2007;@klein_complement_2005].
In the nearly 17 years since publication of early studies, researchers have reported 
thousands of novel SNP-trait associations from GWAS [@klein_complement_2005;@welter_nhgri_2014;@visscher_10_2017;@visscher_five_2012]. Ever larger sample sizes in GWAS have enhanced statistical power to detect associations and refined our understanding of 
human health and disease.

```{r, article-counts, message = FALSE, fig.cap= "Upward trend in article counts by year for keyword searches: 1. mendelian randomization and 2. genome-wide association study."}
library(magrittr)
dat <- readxl::read_xlsx(path = "data/publication-counts.xlsx", sheet = "counts")
dat %>%
  tidyr::pivot_longer(!year) %>%
  dplyr::filter(year < 2022) %>%
  ggplot2::ggplot(ggplot2::aes(x = year, y = value, fill = name)) + 
  ggplot2::geom_bar(stat = "identity", 
                    position=ggplot2::position_dodge()) + 
  ggplot2::scale_fill_manual(labels = c("GWAS", "MR"), values=c('#999999','#E69F00')) + 
  ggplot2::ylab("Annual publication count") + 
  ggplot2::xlab("Year") + 
  #ggplot2::ggtitle("Published journal articles per year for Google Scholar key word searches: 1. Mendelian randomization and 2. Genome-wide association study") + 
  ggplot2::guides(fill = ggplot2::guide_legend("Search terms")) + 
  ggplot2::theme(text = ggplot2::element_text(size = 20))

ggplot2::ggsave("figures/mr-gwas-counts.png")

```

Methods for and uses of MR have appeared at a rapid and accelerating pace since the 
publication of the early GWAS (Figure \@ref(fig:article-counts)). Many studies use the "two-sample" MR design, in which one cohort of subjects has measurements for the exposure, while a second cohort has measurements for a second trait [@pierce_efficient_2013]. Importantly, the SNP instruments are measured in both sets of subjects.

Inverse variance 
weighted MR [@burgess_mendelian_2013], MR-Egger [@bowden_mendelian_2015], weighted 
median estimation [@bowden_consistent_2016], Bayesian weighted MR 
[@zhao_bayesian_2019], robust adjusted profile score [@zhao_statistical_2020], MRMix 
[@qi_mendelian_2019], CAUSE [@morrison_mendelian_2020], and MRAID 
[@yuan_likelihood_2021] are all recent methods developments. These methods differ in 
their approaches to two challenges:

1. instrument selection (from among the available SNPs)  
1. modeling and controlling for horizontal pleiotropy

Many MR methods use a set of pre-selected SNPs as instruments. Typically, these SNPs are selected to 
be statistically independent. This restriction to 
independent SNPs is needed for valid inference in methods like inverse variance weighting MR [@burgess_mendelian_2013]. The independent SNPs may be 
selected through linkage disequilibrium (LD) clumping [@purcell2007plink].

However, LD clumping for choosing instrument SNPs can be problematic for several 
reasons. First, the chosen SNPs may not be causal for the outcome. This can 
happen, for instance, by using a noncausal SNP that is in high LD with a 
causal SNP. Second, because many traits are polygenic or omnigenic, multiple 
causal SNPs may be in high LD together. LD clumping is likely to omit one or more
of the causal SNPs, which can diminish statistical power in MR analysis.

Horizontal pleiotropy occurs when a SNP instrument affects the outcome through at least one pathway that bypasses the exposure variable [@verbanck_detection_2018]. It constitutes a violation of 
the standard MR assumptions and can lead to biased 
causal effect estimates and diminished statistical power. We discuss approaches to 
modeling and accounting for horizontal pleiotropy below.

Recent methods developments have enabled execution of two-sample MR studies with only 
GWAS summary statistics [@burgess2015using]. These strategies are especially pertinent now that there is 
an abundance of freely available GWAS summary data [@buniello2019nhgri]. The 
ability to use published data has contributed to a rapid increase in the number of
two-sample MR methods for GWAS summary statistics.

# Mendelian Randomization with one Instrument

We begin our discussion of Mendelian randomization in GWAS data by considering the simplest case, where 
there is a single genetic marker and a single outcome variable. Mendelian randomization with one 
instrument uses a single genetic marker 
as the proxy ("instrument" or "instrumental variable") for the "exposure", a complex trait, and asks 
whether variation in the genotypes at the genetic marker causes variation in the 
"outcome" variable, which is typically a second complex trait. When the genetic marker 
is a SNP on an autosome, the cohort is partitioned into three 
classes, two homozygote classes and a single heterozygote class, based on genotypes 
for that marker. When the three assumptions of mendelian randomization are satisfied, 
estimation of the causal effect of the exposure on the outcome is possible.

The outcome variable can be binary, a count, a time to event, or a continuous variable. For brevity and simplicity, 
we focus on continuous outcomes before considering other classes of outcome variables in a later section.

## Assumptions of Mendelian Randomization

Three assumptions are central to MR. Satisfying them enables inference of 
causal effects in an observational MR study. In practice, careful evaluation of the assumptions is needed. 
Because assumptions 2 and 3, which involve absence of certain types of confounding, can't be verified, 
practitioners must be cautious when planning MR analyses. Sensitivity analysis is highly recommended as a 
way to assess robustness of estimates in the presence of possible assumption violations.

The first assumption states that the exposure is associated with the outcome. The second requires 
independence, conditional on the exposure and all measured and unmeasured confounders, between the 
genetic marker and the outcome. The last of the three assumptions states that the genetic marker is 
independent of all (measured and unmeasured) confounders of the relationship between exposure and 
outcome. We next discuss these in more detail and consider their evaluation in MR studies [@vanderweele_methodological_2014].

If the first assumption, about the association between the genetic marker and the exposure, holds, 
but this association is weak, then an amplification of biases, such as those due to violations of 
assumptions 2 and 3, may result [@vanderweele_methodological_2014]. @burgess_avoiding_2011, in analyses 
of simulated and real data, demonstrated that bias in causal estimates increases with decreases in 
the strength of association between the SNP and exposure. We later discuss, in the context of 
methods for multiple SNP instruments, how this finding, coupled with the recognition of 
the polygenicity of many complex traits, has motivated methods to work with large numbers of weak SNP instruments [@ye_debiased_2021;@chao_consistent_2005].

The second MR assumption, often termed the exclusion restriction assumption, can also be stated as 
the need for the genetic marker to affect the outcome only through the exposure. In a 
causal diagram, this means that the only path from the marker to the outcome is that 
containing the exposure.

The third assumption ensures that the genetic marker is
independent of all confounders of the exposure-outcome relationship. Like the exclusion restriction
assumption, we can't verify this assumption with observational data, so analysts need to 
use tools like sensitivity analysis to ensure validity of findings.

When the three MR assumptions hold, one may use instrumental variable statistical methods, 
with the genetic marker data, the exposure data, and the outcome data, to estimate and to test the 
causal effect of the exposure on the outcome [@didelez_mendelian_2007;@lawlor_mendelian_2008;@palmer_adjusting_2008].

To understand the origin of the term "Mendelian randomization", it is helpful to 
recall Mendel's inheritance laws [@mendel1865versuche]. Under the assumption that alleles segregate 
randomly from parent to offspring, the offspring genotypes are unlikely to be 
associated with confounders of the exposure-outcome relationship. Additionally,
reverse causation is unanticipated since germ-line genotypes are 
fixed at conception and, thus, precede realization of other observed variables. Reverse causation 
occurs when the outcome at an early time point has a causal effect on the exposure variable. While 
reverse causation is often assumed to not occur in MR study designs, @burgess_are_2021 points out 
several scenarios where this assumption may not hold [@davey_smith_mendelian_2003].

## Statistical models and methods for MR with one instrument

Approaches to MR can be classified into four categories: ratio of coefficients method, two-stage methods, 
likelihood-based methods, and semiparametric methods.

### Ratio of coefficients method

The ratio of coefficients method, also known as the Wald method, estimates the 
causal effect of the exposure X on the outcome Y by using a single SNP instrument [@wald_fitting_1940]. For a continuous outcome, the causal effect estimator, $\hat\gamma_{Wald}$ is defined in Equation \@ref(eq:wald).

\begin{equation}
  \hat\gamma_{Wald} = \frac{\hat\beta_{YZ}}{\hat\beta_{XZ}}
  (\#eq:wald)
\end{equation}

In Equation \@ref(eq:wald), $\hat \beta_{YZ}$ and $\hat \beta_{XZ}$ are the slope 
estimates from the regressions of the outcome and exposure, respectively, on the 
SNP instrument. Since the Wald method requires only the regression coefficients, it can be used with summary data. However, the Wald method doesn't accommodate multiple SNP instruments, which limits its use in the GWAS setting.

### Two-stage methods

A two-stage statistical model involves two regression models [@burgess_review_2017;@sawa_exact_1969]. For a continuous 
outcome, one may perform "two-stage least squares", which involves two linear regressions. 
First, the exposure variable is regressed on the instrument (*i.e.*, the SNP genotypes) (Equation \@ref(eq:lm1)). We denote each subject's exposure variable value as $x_i$ in Equation \@ref(eq:lm1), while $\alpha_0$ denotes an intercept term and $\alpha_1$ is the slope. $z_i$ denotes the SNP instrument genotype for subject $i$, while $\epsilon_i$, the error term, is assumed to be independent among subjects and normally distributed with a shared common variance and mean zero.

The resulting 
fitted values for the exposure variable from Equation \@ref(eq:lm1) are the independent variable in the second linear regression, 
where the outcome is the dependent variable. The causal effect estimate, then, is the 
coefficient $\hat \beta_1$ obtained from the second regression analysis (Equation \@ref(eq:lm2)). Note also that the random errors in Equation \@ref(eq:lm2) are assumed independent of those in Equation \@ref(eq:lm1). Like the $\epsilon_i$ in Equation \@ref(eq:lm1), the $\tau_i$ are assumed independent and identically distributed normal random errors.

\begin{equation}
  x_i = \alpha_0 + \alpha_1 z_i + \epsilon_i
  (\#eq:lm1)
\end{equation}

\begin{equation}
  y_i = \beta_0 + \beta_1 \hat x_i + \tau_i
  (\#eq:lm2)
\end{equation}

Note that the uncertainty in the fitted values from 
the first regression is not considered when performing the second regression.
For this reason, the variance of the coefficient estimator, $Var(\hat \beta_1)$ is 
incorrect in two-stage calculations. This and other observations led researchers to develop 
likelihood-based MR methods.

### Likelihood-based methods

Likelihood-based MR methods, unlike two-stage methods, provide maximum likelihood 
estimates with their many desirable properties [@fisher1925theory]. Limited information maximum likelihood from econometrics 
is the earliest approach for likelihood-based inference in MR [@davidson1993estimation].

Limited information maximum likelihood with a single SNP instrument is modeled with two equations (Equations \@ref(eq:liml1) and \@ref(eq:liml2)), where the random errors follow a bivariate normal distribution [@burgess_review_2017].

\begin{equation}
  x_i = \alpha_0 + \alpha_1 z_{i1} + \epsilon_{Xi}
  (\#eq:liml1)
\end{equation}

\begin{equation}
  y_i = \beta_0 + \beta_1 x_i + \epsilon_{Yi}
  (\#eq:liml2)
\end{equation}

Limited information maximum likelihood is sometimes called the maximum likelihood 
counterpart of two-stage least squares, and it yields the same causal estimate as 
two-stage least squares and the ratio method when used with a single SNP 
instrument. Additionally, the limited information maximum likelihood framework can accommodate $K > 1$ SNP instruments by replacing $\alpha_1 z_{i1}$ with the sum $\sum_{k = 1}^K\alpha_k z_{ik}$.

One may also use Bayesian methods to obtain likelihood-based estimators [@kleibergen_bayesian_2003]. @burgess_improving_2012 examined a model that is similar to that from the limited information maximum likelihood framework (Equation \@ref(eq:b1)). The Bayesian model differs from the limited information maximum likelihood model in that the causal effect parameter, $\beta_1$ represents the effect between the true means for the exposure and the outcome. In the limited information maximum likelihood model, the causal effect is that of the measured effect on the measured outcome.

\begin{eqnarray}
\begin{pmatrix}
X_i \\
Y_i
\end{pmatrix} \sim N_2\left(\begin{pmatrix} \xi_i\\
\eta_i
\end{pmatrix}, \Sigma
\right)\nonumber \\
\xi_i = \alpha_0 + \alpha_1 z_{i1}\nonumber \\
\eta_i = \beta_0 + \beta_1 \xi_i
  (\#eq:b1)
\end{eqnarray}

For each subject $i$, the exposure and outcome values come from a bivariate normal 
distribution. The mean of the exposure distribution is assumed linear in 
the SNP instrument, and the mean of the outcome distribution is linear in the 
mean exposure [@burgess_review_2017;@jones_choice_2012]. 
@burgess_improving_2012 demonstrated that the Bayesian model, with weak 
instruments, outperforms the limited information maximum likelihood model in terms 
of frequentist coverage levels.

### Semiparametric methods

Semiparametric instrumental variable methods, which feature parametric and nonparametric components, 
typically assume a parametric model connecting the exposure and outcome, but don't 
impose distributional assumptions on the errors [@burgess_review_2017]. Compared to 
fully parametric models, semiparametric models often are more robust to model 
misspecification [@burgess_review_2017;@clarke_instrumental_2012]. We follow 
@burgess_review_2017 by discussing three semiparametric strategies, generalized 
method of moments, continuous updating estimator, and G-estimation of structural 
mean models.

Generalized method of moments can be viewed as a more flexible form of two-stage 
least squares that handles heteroscedastic errors and nonlinearity in the two 
regressions [@burgess_review_2017]. Before specifying the model, we introduce 
notation. $\mathbb{E}(Y|\text{do}(X = x))$ is the conditional expectation of $Y$ if 
we forced $X$ to take value $x$ for every subject [@pearl2009causality].

The generalized method of moments equations, with a single SNP instrument, then, can be written as in Equation \@ref(eq:gmm).

\begin{eqnarray}
  \sum_i (y_i - f(x_i;\beta)) = 0\nonumber \\
  \sum_i z_i(y_i - f(x_i;\beta)) = 0
  (\#eq:gmm)
\end{eqnarray}

The GMM estimate is the value of the vector $\beta$ that satisfies Equation \@ref(eq:gmm), where $f(x_i; \beta) = \mathbb{E}(Y|\text{do}(X = x))$. @palmer_instrumental_2011 developed numerical methods for obtaining estimates from Equations \@ref(eq:gmm).

Another semiparametric estimation method is G-estimation of a structural mean model [@burgess_review_2017;@robins_new_1986;@greenland_estimating_2008].  
We follow @burgess_review_2017 by defining the potential outcome $Y(x)$ as the outcome value that we would 
have observed had we set the exposure value $X$ to $x$. For example, $Y(0)$ denotes the observed outcome had 
we set the exposure to zero instead of its observed value $x$. The structural mean model for a continuous 
outcome is displayed in Equation \@ref(eq:smm).

\begin{equation}
\mathbb{E}(Y(x) - Y(0) | X = x, Z = z) = \beta_1 x
  (\#eq:smm)
\end{equation}

The causal effect parameter is $\beta_1$. @burgess_review_2017 derives the estimating equations for 
$\beta_1$, after noting that the conditional expectation, $\mathbb{E}(Y(0)| X = x, Z = z)$ is 
independent of $Z$ and reasoning that the causal effect is that value of $\beta_1$ that yields zero 
covariance between $Z$ and $\mathbb{E}(Y(0)| X = x, Z = z)$ (Equations \@ref(eq:smm1)).

\begin{equation}
\sum_i (z_{ik} - \bar z_{\cdot k})(y_i - \beta_1 x_i) = 0
  (\#eq:smm1)
\end{equation}

Note that $k$ indexes SNP instruments and ranges from 1 to $K$ in Equation \@ref(eq:smm1).

# Mendelian Randomization with multiple SNP instruments

Recognition of complex traits' diverse genetic architectures has fueled development of two-sample MR 
methods with multiple SNP instruments [@hormozdiari_leveraging_2018;@watanabe_global_2019;@pierce_power_2011;@bowden_mendelian_2015]. 
Choosing a subset of the available SNPs to serve as instruments is one issue that arises 
in these studies. The solution to this task differs among published methods. Early methods chose a 
set of independent SNPs [@verbanck_detection_2018;@zhu_causal_2018;@zeng_causal_2019-2;@zeng_causal_2019-1]. However, this approach can be problematic when there are multiple causal 
SNPs in linkage disequilibrium with each other. In this setting, discarding some of the causal SNPs may only capture a small
proportion of trait variance explained by the exposure and lead to a power loss in MR [@burgess_review_2017;@burgess_mendelian_2013;@burgess_combining_2016;@burgess_bias_2011]. In support of
this is the finding that including multiple correlated SNPs in transcriptome-wide association
studies improves power compared to using independent SNPs only [@liu2021multi;@gusev_integrative_2016;@yuan_testing_2020;@zeng_non-parametric_2017].

## Statistical models and methods for MR with multiple instruments

@smith_mendelian_2004 recognized that many genetic variants individually explain a 
small proportion of the variation in a trait, and, thus, sufficient statistical 
power would require sample sizes in the tens of thousands 
[@schatzkin2009mendelian;@burgess_mendelian_2013]. @palmer_using_2012 partially 
resolved this issue by proposing use of multiple SNP instruments in a single MR 
analysis. @pierce_power_2011 reasoned that a single causal estimate derived from 
a collection of causal SNPs would have greater precision than an estimate derived 
from only one SNP [@burgess_mendelian_2013]. At the same time, sharing of GWAS 
summary statistics was encouraged as a best practice [@nature2012asking]. 
@harbord_severity_2013 recognized that one could calculate the Wald ratio estimate of 
the causal effect from GWAS summary statistics, and @burgess_mendelian_2013 devised 
a strategy for determining a single causal effect from summary statistics for a 
collection of SNPs. @burgess_mendelian_2013 termed their method "inverse variance 
weighting" (IVW) because each causal SNP's effect is weighted by the inverse of the variance 
of the ratio estimator, and the overall causal effect is the sum of the weighted SNP
causal effects.

Specifically, @burgess_mendelian_2013 combined causal effect ratio estimates from $M$ independent SNPs by using Equation \@ref(eq:ivw).

\begin{equation}
  \hat \beta_{IVW} = \frac{\sum_{j = 1}^M \hat\theta_j [Var(\hat\theta_j)]^{-1}}{\sum_{j = 1}^M [Var(\hat\theta_j)]^{-1}}
  (\#eq:ivw)
\end{equation}

In efforts to detect the presence of horizontal pleiotropy in two-sample MR studies with GWAS summary data, @greco_m_detecting_2015 developed the between-instrument heterogeneity Q test. In a meta-analysis of MR Wald ratio estimates, one per SNP instrument, they calculate Q in Equation \@ref(eq:q).

\begin{equation}
Q = \sum_{k = 1}^K w_k(\hat\beta^{(k)}_{XY} - \mu_F)^2
(\#eq:q)
\end{equation}

In Equation \@ref(eq:q), $w_k$ is the inverse variance of the Wald estimator, and $\mu_F = \frac{\sum_{k = 1}^K w_k \hat\beta^{(k)}_{XY}}{\sum_{k = 1}^K w_k}$.

While the Q test tends to be conservative in small sample sizes, its power increases with increasing 
sample size and increasing degree of pleiotropy [@greco_m_detecting_2015].

@bowden_mendelian_2015 adapted Egger regression, originally developed 
to detect bias in meta-analyses, to detect bias from horizontal 
pleiotropy in two-sample MR studies with GWAS summary data 
[@egger1997bias]. The slope coefficient estimates the causal effect. 
Under the assumption that the SNP-exposure relationship is independent 
of the SNP's pleiotropic effects, Egger's test offers a valid test of 
the null causal hypothesis and consistently estimates the causal 
effect, even if all SNPs are invalid instruments [@bowden_mendelian_2015].

@bowden_consistent_2016, noting that consistency with IVW MR is not robust to 
invalid instrument use, developed a method for consistent MR analysis with  
weighted median-based estimators, building on research from @han2008detecting. 
Its breakdown point is 50\%, 
meaning that up to 50\% of SNP instruments can be invalid while maintaining consistency in estimation 
[@hampel1971general]. Specifically, let $\tilde \beta_j$ denote the $j^{th}$ ordered ratio estimate 
statistic, from least to greatest. If an odd number, say $2K +1$, of SNP instruments is used, then 
the simple median estimator is defined as the $K + 1$ ordered ratio estimate. For an even number, say 
$2K$, number of SNP instruments, choose the midpoint between the two middle-ranking ordered ratio 
estimates, $\frac{\tilde \beta_K + \tilde \beta_{K + 1}}{2}$.

Due to the inefficiency of the simple median estimator, @bowden_consistent_2016 define the weighted 
median estimator. To construct this estimator, order the ratio estimates from least to greatest, as 
with the simple median estimator. The weighted median estimator, then is the median of a distribution
defined by having $\tilde \beta_j$ as the $p_j = 100(\sum_{k = 1}^j - \frac{w_j}{2})$ percentile, 
where $w_j$ is the weight assigned to the $j^{th}$ ordered ratio estimate. In this setting, the 
simple median estimator is seen to be the weighted median estimator when all weights are equal.

@bowden_consistent_2016 also studied the penalized weighted median estimator. They defined the 
penalized weights, $w_j^* = w_j' \times \min(1, 20q_j)$, where $q_j$ is the p-value resulting from 
the comparison of $Q_j = w_j'(\tilde\beta_j - \hat\beta_{IVW})^2$ to 
a $\chi^2_1$ distribution [@bowden_consistent_2016;@cochran1950comparison]. The penalized 
weighted median estimator leaves most variants unaffected, but downweights those with outlying Wald 
ratio estimates.

@kang_instrumental_2016 developed a LASSO-based method, sisVIVE, to 
identify invalid SNP instruments [@tibshirani1996regression]. sisVIVE's 
advantage over earlier methods is that it doesn't require the analyst 
to know which SNP instruments are valid. Instead, it requires that at 
least 50\% of the instruments be valid. sisVIVE outperforms two-stage 
least squares in many ways and performs similar to oracle two-stage 
least squares. Simulations and data analysis results reveal that 
sisVIVE is robust to possibly invalid instruments.

@windmeijer_use_2019, building on the research from 
@kang_instrumental_2016, implemented an adaptive LASSO-based estimator 
after recognizing that sisVIVE misclassifies valid SNP instruments as 
invalid when the invalid SNP instruments have strong effects on the 
exposure. Consistent selection of invalid SNP instruments, they found, 
depends on SNP instrument correlations. To address this issue, @windmeijer_use_2019 
proposed a median estimator with consistency that doesn't 
depend on SNP instrument-exposure association strength or the SNP instrument 
correlation structure. They then applied methods from @zou2006adaptive to 
achieve a consistent estimator with the same asymptotic distribution as the oracle 
two-stages least squares. One important limitation of this thread of research is that
both @kang_instrumental_2016 and @windmeijer_use_2019 require individual-level data.

@hartwig_robust_2017 developed a weighted mode-based MR causal effect estimation method for two-sample MR analysis with GWAS summary data. 
They first calculate Wald ratio estimates for every instrumental SNP. They then apply 
smoothing to the empirical distribution of ratio estimates. The mode of this smoothed 
distribution is the simple mode-based estimate of the causal effect 
[@hartwig_robust_2017]. The inverse variance-weighted mode-based estimate is obtained by 
weighting the empirical distribution of ratio estimates by the inverse variance of each 
estimate. @hartwig_robust_2017 compared their weighted and unweighted mode-based 
estimators with weighted and unweighted median-based estimators, MR Egger regression, and 
IVW. They found that their mode-based estimators demonstrated less bias and lower type I error rates than 
other estimators in some simulation settings. The mode-based estimators possessed lower statistical power to 
detect a causal effect compared to the IVW and weighted median-based methods, but their power exceeded that of 
MR Egger regression [@hartwig_robust_2017]. The mode-based estimators consistently estimated the causal effect 
when the mode across SNP instruments of the horizontal pleiotropy effects was zero [@hartwig_robust_2017]. In 
this manner, the mode-based estimators demonstrated a greater robustness to horizontal pleiotropy than did the 
other estimators.

@rees_robust_2019 approached the challenge of horizontal pleiotropy by 
developing three new methods for two-sample MR analysis with GWAS summary data: robust 
regression, penalized weights, and LASSO penalization. The first two can be viewed as 
modifications of MR-Egger and IVW methods. Together, they offer three strategies for 
downweighting or excluding variants with heterogeneous causal estimates.
Recall that MR-Egger offers consistent causal effect estimates when there is no 
horizontal pleiotropy or when the horizontal pleiotropy effects adhere to the 
"Instrument Strength Independent of Direct Effect" (InSIDE) assumption [@bowden_mendelian_2015]. The 
InSIDE assumption is satisfied when there is no correlation between the pleiotropic SNP 
instrument-outcome effects and the SNP instrument-exposure effects. @rees_robust_2019 uses 
MM estimation, which is one technique for robust linear regression [@koller2011sharpening], to 
formulate an estimator with high breakdown point and high efficiency. The second estimation 
method, with penalized weights, is much like the penalized median estimator, from above, 
except that the second argument in the minimum function is $100q_j$, instead of the $20q_j$ 
above. Finally the LASSO-based method fashions an estimator by applying a penalty to the 
objective function from MR-Egger to get Equation \@ref(eq:lasso) as the needed objective function.

\begin{equation}
\sum_j [Var(\hat\beta_j)]^{-1}(\hat\beta_{Yj} - \theta_{0j} - \theta_1 \hat\beta_{1j})^2 + \lambda \sum_j |\theta_{0j}|
(\#eq:lasso)
\end{equation}

@rees_robust_2019 studied two procedures for choosing $\lambda$: a heterogeneity stopping rule 
and a cross-validation rule. Taken together, these three methods offer additional contributions to 
a suite of sensitivity analysis tools that practitioners should consider.

@burgess_robust_2020 introduced the contamination mixture method for robust and efficient MR
analysis with hundreds of SNPs, some of which may be invalid instruments by violating one or 
more of the three core MR assumptions. Working in the setting of two-sample MR with GWAS data, they begin by 
considering a collection of candidate SNP instruments, which may contain some invalid SNP instruments. For 
each candidate SNP instrument, they calculate the Wald ratio estimate of the causal effect. With the 
collection of Wald ratio estimates, @burgess_robust_2020 reason that the set of valid SNP instruments have 
Wald ratio estimates that arise from a normal distribution centered on the true causal parameter with variance 
equal to that of the Wald ratio estimator. The invalid instruments, however, are assumed to follow a normal 
distribution centered at zero with a variance greater than that of the Wald ratio estimator. They then work 
with profile log likelihood over a grid of causal parameter values to make inferences.

@zhao_statistical_2020 presented MR-RAPS, a method that leverages a robust adjusted 
profile score to accomplish statistical inference in two-sample MR analyses with GWAS 
summary data. Dividing pleiotropic effects into systematic and idiosyncratic, they model the 
systematic pleiotropy with random effects. In so doing, no SNP satisfies the exclusion 
restriction assumption. Adjusting their profile likelihood estimator from the setting 
without pleiotropy, @zhao_statistical_2020 obtain an estimator with consistency and 
asymptotic normality [@mccullagh_simple_1990]. Idiosyncratic pleiotropy is addressed through robustification of 
the adjusted profile score [@huber_robust_1964]. They then demonstrate the properties of their estimator by 
analyzing simulated and real data [@zhao_statistical_2020].

@bowden_improving_2019, working with a collection of SNP instruments, reasoned that if all three MR assumptions are satisfied, then the 
collection of Wald ratio estimates should be homogeneous; thus, any departures from 
homogeneity may signal a violation of one or more assumptions. They reported 
modified weights, in addition to studying weights derived from first order and 
second order approximations to the variance. Their modified weights lead 
to an estimator that is useful for quantifying heterogeneity and detecting outliers.

@wang_causal_2021 developed a framework, GRAPPLE, for performing two-sample MR with weak and strong 
instruments. Using GWAS summary statistics as inputs, GRAPPLE extends MR-RAPS and can detect presence of horizontally 
pleiotropic pathways, infer the causal direction, and perform multivariable MR. Central to the 
GRAPPLE framework is the observation that a horizontally pleiotropic pathway often gives rise to an 
additional mode in the profile likelihood. GRAPPLE uses the presence of multiple profile likelihood 
modes to diagnose horizontal pleiotropy with effects that are grouped by pathway. 
GRAPPLE facilitates identification of SNP instruments that contribute to the 
additional profile likelihood modes. A researcher using GRAPPLE may then recognize 
a confounding factor for each mode. With additional GWAS summary data with traits 
that reflect the confounding factors, GRAPPLE can fit multivariable MR models when
the InSIDE assumption holds for the remaining horizontally pleiotropic effects [@wang_causal_2021].

@foley_mr-clust_2021 recognized that in some MR studies, the Wald ratio estimates for different
SNP instruments may form clusters, and these clusters may represent distinct causal mechanisms 
and may identify the SNPs that are involved in them. They devised a method to detect 
these clusters in the collection of Wald ratio estimates and found SNP instruments that 
point to distinct causal pathways of biological importance [@foley_mr-clust_2021].

MR-GENIUS is a framework for two-sample MR analysis when individual-level data are available [@tchetgen_tchetgen_genius_2021]. @tchetgen_tchetgen_genius_2021 focused efforts on relaxing the 
exclusion restriction assumption and building on existing G-estimation 
methods to create an estimator that is robust to violations 
of the exclusion restriction and to additive unmeasured confounding [@robins1994correcting]. They observe that their estimator, in some settings, reduces to that of @lewbel2012using, which is widely used in econometrics, but not in the MR literature.

@fang2021evaluating

@sadreev2021navigating

### MR with allele scores

Another approach to integrating multiple SNP instruments into a MR framework is
through calculation of allele scores [@burgess_use_2013]. @burgess_use_2013 
calculate an unweighted score as the number of exposure-increasing alleles in 
the subject's genotypes. They also calculate a weighted allele score by using 
the exposure effects as weights. For example, a subject with $g_k$ copies of 
exposure-increasing alleles for SNP $k$ has an unweighted allele score $z = \sum_{k = 1}^K g_{k}$ 
and a weighted allele score $z = \sum_{k = 1}^K w_kg_k$.

@burgess_use_2013 found that when weights are obtained from external data or from 
cross-validation or jackknife approaches applied to the analysis data, the allele score 
functions as a single instrumental variable and greatly diminishes the bias compared to that of the 
two-stage least squares estimator [@burgess_review_2017;@angrist1999jackknife].

# Assumption violations

It is worthwhile to consider ways in which the standard MR assumptions may be violated. 
Let's first consider the exclusion restriction assumption (assumption 2 above). It states that the genetic marker 
must affect the outcome only through the exposure. The causal diagram in Figure \@ref(fig:dag1) illustrates 
a scenario that violates this assumption when the variable X is used as the exposure [@vanderweele_methodological_2014]. Bias in causal effect estimation would result. However, 
if variable W is measured and treated as the exposure, then an unbiased estimate of causal effect is 
possible, since W captures the two pathways, W -> Y and W -> X -> Y, through which the genetic marker G 
affects the outcome Y [@vanderweele_methodological_2014;@textor2011dagitty;@textor2016robust].

```{r, dag1, fig.cap = "Scenario where the genetic variant G affects an intermediate variable W on the pathway to the exposure X. Because W affects Y through a pathway that doesn't involve exposure X, this scenario violates the exclusion restriction assumption."}

library(dagitty)
library(ggplot2)

d1 <- dagitty('dag {
bb="0,0,1,1"
G [pos="0,0"]
W [pos="0.5,0"]
X [exposure, pos="1.0,0"]
Y [outcome, pos="1.5,0"]
G -> W
W -> X
W -> Y [pos="1,1"]
X -> Y
}')

plot(d1)

```

```{r, dag2, fig.cap = "Scenario where the genetic variant G affects exposure X, which affects Y. Unobserved confounder U affects both X and Y."}
d1 <- dagitty('dag {
bb="0,0,1,1"
G [pos="0.0,0.0"]
U [pos="0.75,0.5"]
X [exposure, pos="0.5,0.0"]
Y [outcome, pos="1.0,0.0"]
G -> X
U -> X
U -> Y
X -> Y
}'
)
plot(d1)

```

Recognition of the many ways that the MR assumptions may be violated has inspired methods advances that 
enable assumption relaxations and has motivated use of sensitivity analysis to quantify the impact of 
possible assumption violations.

# Sensitivity analysis

Sensitivity analysis is recommended and widely used in MR studies because of the inability to verify 
the three MR assumptions with observational data. The goal of a sensitivity analysis is to gain insight 
into how the results might differ if the assumptions be violated.
Because not all confounding 
variables are known or measured, assumptions 2 and 3 are not fully verifiable 
in MR studies. Recognition of this fact has inspired the development of 
sensitivity analysis tools. While we can't assess whether a SNP instrument is 
associated with every confounder of the exposure-outcome association, it is 
possible to examine the associations of the SNP instrument with the measured 
covariates. While absence of such associations doesn't guarantee satisfaction 
of the assumption, presence of SNP-covariate associations must be investigated
carefully, as they may constitute assumption violations [@burgess_sensitivity_2017].

@burgess_sensitivity_2017 considered a collection of sensitivity analysis methods when 
working with GWAS summary statistics. @burgess_sensitivity_2017 presented methods for both assessing the
MR assumptions, to the extent possible, and performing robust analyses. For example, 
they used measured covariates to assess for possible associations with the SNP 
instruments. While they can't rule out the possibility of unmeasured confounding, they can study the possibility of measured covariates serving as confounders.

To illustrate their approach, @burgess_sensitivity_2017 shared a case study in which they examined the causal effect of 
C-reactive protein (CRP) levels on coronary artery disease risk with four genetic 
variants in the CRP gene region and 17 other genetic variants that affect coronary 
artery disease risk. They used measured covariates to probe for SNP-covariate 
associations, and followed it up with scatter plots and Cochran's Q test on the causal 
estimates to inquire about whether the 
SNP instruments all identify the same causal parameter [@greco_m_detecting_2015;@small_sensitivity_2007;@higgins_measuring_2003]. 
Additionally, @burgess_sensitivity_2017 suggested using a funnel plot, like those in the 
meta-analysis literature, to visualize possible evidence of directional pleiotropy, where the average pleiotropic effects of the SNP instruments is nonzero [@sterne_recommendations_2011]. Additionally, Egger regression can be useful in this 
setting [@egger_bias_1997;@bowden_mendelian_2015;@burgess_interpreting_2017;@bowden_framework_2017].

# Horizontal pleiotropy

Pleiotropy, where a single genetic variant affects multiple traits, has a long history of study in genetics and complex traits [@solovieff_pleiotropy_2013;@visscher_plethora_2016;@stearns_one_2010]. 
Horizontal pleiotropy refers to the presence of multiple causal pathways from a genetic 
variant to the outcome.
The fact that there are pathways that exclude the exposure and connect the SNP to the outcome violates
standard MR assumptions. @verbanck_detection_2018 documented widespread horizontal pleiotropy 
in GWAS. This finding motivated the development of MR-PRESSO and inspired 
further research into methods to model horizontal pleiotropy [@verbanck_detection_2018].

In the short time since publication of @verbanck_detection_2018, researchers have recognized two 
types of horizontal pleiotropy [@yuan_likelihood_2021]. The first occurs via 
exposure-independent paths. The resulting horizontal pleiotropic effects are independent of the
SNP-exposure relationships. The second type of horizontal pleiotropy manifests
in the presence of unobserved exposure-outcome confounding. It induces correlation between 
horizontal pleiotropic effects and SNP-exposure effects. Both types of horizontal pleiotropy 
violate standard MR modeling assumptions and can bias causal effect estimates and can increase false 
discoveries [@yuan_likelihood_2021].
Early MR analyses avoided confounding from horizontal pleiotropy by discarding 
instrumental SNPs that might be associated with the outcome 
[@verbanck_detection_2018;@zhu_causal_2018;@zeng_causal_2019-1;@zeng_causal_2019-2].
More recent methods have attempted to model horizontal pleiotropy 
[@morrison_mendelian_2020;@yuan_testing_2020;@yuan_likelihood_2021]. CAUSE [@morrison_mendelian_2020] and 
MRMix [@qi_mendelian_2019] both use a mixture of normal distributions to control for both 
types of horizontal pleiotropy. Modeling both types of horizontal pleiotropy is particularly 
challenging because the MR model likelihood often involves an integral that can't be solved analytically. 
Because of this issue with the model likelihood, both MRMix and CAUSE use other, non-likelihood-based methods for inference.

@hemani_orienting_2017

# Correlated horizontal pleiotropy

The second type of horizontal pleiotropy, from above, is sometimes called 
"correlated horizontal pleiotropy". Correlated horizontal pleiotropy occurs 
when a SNP affects both the exposure and a variable that confounds the 
relationship between exposure and outcome (Figure \@ref(fig:dag3)).

```{r, dag3, fig.cap = "Correlated horizontal pleiotropy occurs when the genetic variant G affects exposure X, which in turn affects Y, and G affects U, which in turn affects X and Y independently."}
d1 <- dagitty('dag {
bb="0,0,1,1"
G [pos="0.0,0.0"]
U [pos="0.75,0.5"]
X [exposure, pos="0.5,0.0"]
Y [outcome, pos="1.0,0.0"]
G -> X
G -> U
U -> X
U -> Y
X -> Y
}'
)
plot(d1)

```

CAUSE and MRAID are two methods that aim to model correlated horizontal pleiotropy [@morrison_mendelian_2020;@yuan_likelihood_2021]. 
Their approaches to modeling correlated horizontal pleiotropy differ in important ways that ultimately 
affect performance. 
CAUSE allows for a proportion of SNPs to exhibit correlated pleiotropy, which @morrison_mendelian_2020 
models as an effect on a shared, unobserved factor. The remaining SNPs are independent of this 
unobserved factor. Every SNP can have a nonzero pleiotropic effect on the outcome, and these 
pleiotropic effects are uncorrelated with the variant effects on the exposure. The model, then, 
is written as a mixture of biavariate normal distributions, and inference proceeds by borrowing ideas on 
adaptive shrinkage from @stephens_false_2016. Finally, @morrison_mendelian_2020 compares two models 
to determine whether the GWAS summary statistics are consistent with a causal effect of the        
exposure on the outcome. Specifically, they estimate the difference in the expected log pointwise 
posterior density for the model with causal effect fixed at zero and the model that permits a nonzero 
causal effect [@vehtari_practical_2017].

@yuan_likelihood_2021 found that MRMix is not robust to misspecification of SNP effect sizes and often is 
biased. CAUSE, they found, yields overly conservative p-values [@yuan_likelihood_2021].
These observations motivated @yuan_likelihood_2021 to develop a new method, MRAID. MRAID accommodates both individual-level data and GWAS summary statistics [@yuan_likelihood_2021]. We focus on 
the approach that uses GWAS summary statistics. Two equations are central to MRAID modeling (Equations \@ref(eq:mraid1) and \@ref(eq:mraid2)).

\begin{equation}
  \hat{\beta_X} = \Sigma_1\beta + e_x
  (\#eq:mraid1)
\end{equation}

\begin{equation}
  \hat{\beta_Y} = \alpha \Sigma_2 \beta + \Sigma_2\eta_0 + \Sigma_2 \eta_1 + e_y
  (\#eq:mraid2)
\end{equation}

The estimated marginal effects for the $p$ SNP instruments on the exposure are denoted by $\hat \beta_X$, while 
$\hat\beta_Y$ represents the estimated effects of the same SNPs on the outcome. The $p$ by $p$ SNP instrument correlation matrices for the exposure and outcome 
are written as $\Sigma_1$ and $\Sigma_2$, respectively. $\Sigma_1$ and $\Sigma_2$ can be estimated with 1000 Genomes 
Project data, for example, by choosing a subset of 1000 Genomes Project subjects that have similar ancestry 
[@the_1000_genomes_project_consortium_1000_2012]. The $p$-vector error terms, $e_x$ and $e_y$, follow multivariate normal 
distributions with mean zero and variances $\Sigma_1\frac{\sigma_X^2}{n_1 - 1}$ 
and $\Sigma_2\frac{\sigma_Y^2}{n_2 - 1}$, 
respectively. @yuan_likelihood_2021 then constructs a Gibbs sampler to make likelihood-based inferences [@geman1984stochastic]. To facilitate computations, the investigators make several assumptions about the collection of SNPs:

1. Relatively small proportion of SNPs have nonzero effects on the exposure  
1. A relatively small proportion of SNPs demonstrate horizontal pleiotropy  
1. The chosen instrumental SNPs are more likely to display horizontal pleiotropy than are non-instrumental SNPs  
1. Those SNPs that display horizontal pleiotropy are more likely to demonstrate uncorrelated horizontal pleiotropy than correlated horizontal pleiotropy.

By encoding these assumptions in prior distributions, @yuan_likelihood_2021 enables the inferential procedures to accommodate the observed data in the context of the assumptions on SNPs.

\begin{eqnarray}
U = X\beta_U \odot Z_U + \epsilon_U\nonumber \\
A = U\nu + B\delta + X\beta_A \odot Z_A + \epsilon_A \nonumber \\
B = U\eta + A\gamma + X\beta_B \odot Z_B + \epsilon_B 
  (\#eq:wwer)
\end{eqnarray}

To develop MR methods that are robust to correlated horizontal pleiotropy, 
@brown_welch-weighted_2021 present Welch-weighted Egger regression. Their flexible model 
accommodates both uni-directional and bi-directional effects between exposure and outcome. 
@brown_welch-weighted_2021 specifies their model via Equations \@ref(eq:wwer). They denote 
the two traits by $A$ and $B$, the SNP genotypes matrix by $X$, and the unobserved confounder by $U$. $Z_U$, $Z_A$ and $Z_B$ denote the indicator vectors of zeros and ones to indicate which . Note that the model is agnostic with regards to which is the 
outcome and which the exposure, so we follow the authors' notation here.

In Equations \@ref(eq:wwer),

# Binary, Count, and Time-to-event outcomes and exposures

While we've focused above on continuous outcomes and continuous exposures, some
data are more naturally treated as binary, count, or time-to-event variables. MR methods 
for non-continuous outcomes and exposures is an active area of research. @burgess_improving_2012 and @allman_novel_2021 
developed MR strategies for binary exposures and binary outcomes. @burgess_improving_2012 extended 
methods from @wald_fitting_1940 by fitting two logistic regressions, one for the outcome and one for 
the exposure. They then calculated the causal effect estimator as the ratio of the two logistic 
regression coefficient estimators. @allman_novel_2021 devised a new MR method for binary outcome and 
binary exposure by drawing on @gauvreau_analysis_1997 and treating the exposure and outcome as correlated binary 
random variables. They implemented an iterative optimization algorithm to 
simultaneously infer the causal effect parameter and other parameters.

Researchers have also leveraged other generalized linear models for analysis in MR 
studies [@nelder1972generalized]. However, this area is relatively unexplored. @hazewinkel_mendelian_2022 treated hospitalizations as a count variable and 
modeled it with quasi-Poisson methods. @allman_mendelian_2022 characterized statistical properties, including bias and 
power, with simulations of count random variables and binary random variables. With growing interest
in modeling counts in biomedical data, where molecular phenotyping technologies now acquire RNA molecule counts and protein abundances, 
there may be opportunities for methods innovations in a generalized linear models framework for MR studies [@hwang_single-cell_2018;@ozsolak_rna_2011;@domon_mass_2006;@bantscheff_quantitative_2007].

Recently, multiple teams of investigators have analyzed time-to-event traits in MR studies [@burgess_commentary_2015;@arvanitis_linear_2021;@tikkanen_body_2019;@he_causal_2018;@guo_body_2017]. 
@guo_body_2017, for example, studied post-diagnosis survival time in women with breast cancer. They 
treated survival as a time-to-event outcome, which enabled them to model it with a proportional 
hazards model [@cox_regression_1972] in two-stage MR with individual-level data. Similarly, 
@he_causal_2018, in a study of cardiovascular disease, applied proportional hazards regression when 
examining time to stroke as an outcome variable. In very recent work, @tchetgen_tchetgen_genius_2021
reported methods, MR GENIUS, that are robust to violations of the exclusion restriction assumption (under 
many assumed MR data generating processes). They considered binary, continuous, and time-to-event 
outcomes within their framework. We are unaware of MR studies that treat 
the exposure variable as a time-to-event, but it's conceivable that future research will consider 
time-to-event traits as exposures in MR analyses.

# Recent Findings

## Omnigenic MR

Recent research on the omnigenic hypothesis, which posits 
that every SNP's effect on a trait is nonzero, has informed
MR methods development [@Boyle2017]. @wang_mendelian_nodate 
recognized the limitations of previous MR methods in the 
context of the omnigenic hypothesis. In response, @wang_mendelian_nodate 
developed a MR method that uses all genome-wide SNPs as instruments. 
Using GWAS summary statistics as inputs, their method relies on a composite likelihood framework for scalable computation and allows for horizontal pleiotropy. @wang_mendelian_nodate used extensive 
simulations, including those with model misspecifications, to characterize their method's statistical power and robustness. Finally, they applied the new method to 
identify multiple complex traits that affect coronary artery disease and asthma. These causal relationships highlight the important roles of plasma lipids, blood pressure, and the immune system in CAD susceptibility and those of obesity and the immune system in asthma development.

## Multivariate exposures in MR analysis

@grant_pleiotropy_2021: we introduce three methods which build on common approaches in the univariable setting: {MVMR}-Robust; {MVMR}-Median; and {MVMR}-Lasso

# Applications of MR in GWAS

MR analysis has revealed mechanistic insights for many 
common diseases [@mancuso2017integrating;@parisinos2020genome;@zeng_causal_2019-2;@larsson2019smoking;@zheng2020phenome;@robinson2016insight;@luo2019association;@au2018impact;@zhao2019effect;@zeng_causal_2019;@zeng_causal_2019-1]. In some cases, a randomized clinical trial is not possible for ethical or other reasons. 
Below, we discuss recent applications of two-sample MR to GWAS summary statistics.

Birthweight is thought to impact long-term susceptibility to adulthood diseases. This scientific 
hypothesis, sometimes called the "fetal origins of adult diseases" motivated work by @zeng_causal_2019 in 
which the authors used MR to identify causal relationships between lower birthweight and greater risks of 
coronary artery disease, myocardial infarction, and type 2 diabetes mellitus [@barker_fetal_1990;@lucas_fetal_1999;@barker_fetal_2002]. Specifically, 
@zeng_causal_2019 used 
inverse variance-weighted MR to examine a total of 21 adult diseases and 38 other complex traits. Three causal associations met the Bonferroni multiple hypothesis testing threshold 
[@bonferroni1936teoria]. The authors analyzed data from 37 GWAS with up to 340,000 subjects with European 
ancestry [@zeng_causal_2019]. They estimated odds ratios of 1.34, 1.30, and 1.41 for coronary artery 
disease, myocardial infarction, and type 2 diabetes mellitus, respectively. Sensitivity analyses revealed 
the robustness of these findings to possible MR assumptions violations, including pleiotropy, maternal 
effects, and reverse causation [@zeng_causal_2019]. Curiously, none of the 38 other complex traits 
appeared to mediate the causal associations, perhaps due to limited sample sizes for these traits [@zeng_causal_2019]. @zeng_causal_2019 argue that further 
studies, including consideration of socioeconomic status and incorporation of larger sample sizes, may 
reveal mediating factors for these causal relationships and, thus, provide biological and public health insights.

Amyotrophic lateral sclerosis is an incurable and largely untreatable neurological condition [@rowland_amyotrophic_2001;@kiernan_amyotrophic_2011]. Working with GWAS summary statistics, @zeng_causal_2019-1 found evidence of causal relationships between 
type 2 diabetes mellitus and amyotrophic lateral sclerosis in both European and East Asian populations. 
Interestingly, these effects differ in direction; in the European population, type 2 diabetes mellitus 
exerted a neuroprotective effect and diminished the risk of amyotrophic lateral sclerosis, while in the 
East Asian population type 2 diabetes mellitus increased the risk of amyotrophic lateral sclerosis. 
@zeng_causal_2019-1 used inverse variance-weighted methods to estimate the causal effect from summary 
statistics derived from approximately 660,000 type 2 diabetes mellitus European subjects and 81,000 
amyotrophic lateral sclerosis European subjects. Smaller sample sizes (approximately 191,000 type 2 
diabetes mellitus subjects and 4100 amyotrophic lateral sclerosis subjects) were available in the East 
Asian cohorts. The authors used multiple SNP instruments with strong associations with type 2 diabetes 
mellitus [@zeng_causal_2019-1]. The odds ratios of 0.93 and 1.28 in the Europeans and East Asians, 
respectively, were robust against instrument selection, various model misspecifications, and estimation 
biases [@zeng_causal_2019-1]. Egger regression and MR-PRESSO analyses were consistent with the absence of horizonal pleiotropy 
[@zeng_causal_2019-1;@hemani_evaluating_2018;@verbanck_detection_2018;@ong2019implementing]. Future studies may reveal deeper biological insights into these observations and, ultimately, 
contribute to treatments for amyotrophic lateral sclerosis.

Birth weight is associated with childhood asthma, but whether there is a causal relationship 
between the two remains unclear [@brooks_impact_2001]. To investigate this possibility, 
@zeng_birth_2019 performed a MR analysis to assess the causal effect of birth weight on adult 
asthma. They used data from two GWAS, with 140,000 subjects in the birth weight study and 62,000 
subjects in the adult asthma study. With both the genetic risk score method and the 
inverse variance-weighted method, @zeng_birth_2019 found no evidence of a causal effect of birth
weight on adult asthma. Sensitivity analysis results offered reassurance that the 
finding of no causal effect is robust. MR-PRESSO and MR-Egger found no evidence for bias from 
SNP instrument outliers or horizontal pleiotropy. @zeng_birth_2019 concludes that confounding
in other studies may explain the discrepancy between their findings and others' conclusions.

# Summary and Outlook

Modern GWAS for biobank-scale data offer the possibility
of discovering even more causal relationships among complex traits. The large
sample sizes in biobank studies, often greater than 100,000 
subjects, offer the possibility of uncovering associations of modest strength 
that might be undetectable in smaller studies. For example, the UK 
Biobank Study involves more than 500,000 subjects. Each UK Biobank subject's 
data consists of millions of SNP genotypes and measurements for 
thousands of traits [@bycroft_uk_2018].

# References