### 2022-03-04 omitted.Rmd
started omitted.Rmd for removed refs

---
title: "Omitted lit"
author: "Frederick Boehm"
date: "3/3/2022"
bibliography: exported.bib
output: html_document
---

Genetic instrumental variable regression uses GWAS summary statistics to construct polygenic scores for 
the outcome of interest [@diprete_genetic_2018].

# references