### 2022-03-04 omitted.Rmd

---
title: "Omitted lit"
author: "Frederick Boehm"
date: "3/3/2022"
bibliography: exported.bib
output: html_document
---

Genetic instrumental variable regression uses GWAS summary statistics to construct polygenic scores for 
the outcome of interest [@diprete_genetic_2018].

@hemani_automating_2017

@howey_bayesian_2020 use anchored Bayesian networks to

@jiang_hierarchical_2021, building on research from @newcombe2016jam, present a two-stage hierarchical Bayesian 
model that aims to unify MR and transcriptome-wide association studies frameworks.

@mounier_bias_2021
@hartwig_two-sample_2016
@lawlor_commentary_2016
@burgess_bias_2016

@darrous_simultaneous_2021

\begin{eqnarray}
U = X\beta_U \odot Z_U + \epsilon_U\nonumber \\
A = U\nu + B\delta + X\beta_A \odot Z_A + \epsilon_A \nonumber \\
B = U\eta + A\gamma + X\beta_B \odot Z_B + \epsilon_B 
  (\#eq:wwer)
\end{eqnarray}

To develop MR methods that are robust to correlated horizontal pleiotropy, 
@brown_welch-weighted_2021 present Welch-weighted Egger regression. Their flexible model 
accommodates both uni-directional and bi-directional effects between exposure and outcome. 
@brown_welch-weighted_2021 specifies their model via Equations \@ref(eq:wwer). They denote 
the two traits by $A$ and $B$, the SNP genotypes matrix by $X$, and the unobserved confounder by $U$. $Z_U$, $Z_A$ and $Z_B$ denote the indicator vectors of zeros and ones to designate which . Note that the model is agnostic with regards to which is the 
outcome and which the exposure, so we follow the authors' notation here. The symbol $\odot$ denotes the Hadamard elementwise product.

In Equations \@ref(eq:wwer),

## Multivariate exposures in MR analysis

@grant_pleiotropy_2021: we introduce three methods which build on common approaches in the univariable setting: {MVMR}-Robust; {MVMR}-Median; and {MVMR}-Lasso

# references